---
title: "ANALYSIS OF MONTHLY CRUDE OIL PRICE"
author: Jung-Han Wang, Qing Cai, Yina Wu
output: 
  revealjs::revealjs_presentation:
    theme: league
    center: TRUE
---

```{r,echo=FALSE, include=FALSE}
## Set up knitting rule
knitr::opts_chunk$set(echo = TRUE, fig.width=6, fig.height=3,warning=FALSE,message = FALSE)
library(pander)
```

## Objective

Analyze and forecast crude oil prices using the data from West Texas Intermediate (WTI)

## {data-background-iframe="https://fred.stlouisfed.org/series/DCOILWTICO/downloaddata"}

##  Data Preparation

<section> 
Using _data.table_ package to efficiently load csv into R</h2>
```{r}
library(data.table)
gas<-fread("https://www.dropbox.com/s/38awxzkkz1ksmyp/Raw%20Data.csv?raw=1")
```

</section>

<section>
<h4>Check the first row of the data. And identify its class.</h4>
```{r}
head(gas)[1,]
## Check the format
pandoc.table(apply(gas,2,class))
```
</section>

<section>

- Change DATE format from character to Date 
- Changed gas price from character to numeric
- Check all formats are correct

```{r}
### Change Date format to Date
gas$DATE<-as.Date(gas$DATE)
### Change Value format to Numeric
gas$VALUE<-as.numeric(as.character(gas$VALUE))
## Check all format which looks good, DATE=Date, VALUE=numeric
str(gas)
```
</section>

<section>

- Remove missing value in gas price
- Calculate monthly average gas price based on daily price

```{r}
## Remove missing value in the data
gas<-subset(gas,!{is.na(VALUE)})
```

```{r}
## Get the monthly average
gasMonth<-aggregate(list(monthly_value = gas$VALUE), 
                    list(month = cut(gas$DATE, breaks="1 month")), 
                    mean)
```

</section>

<section>
- Convert the data into time series object using function __ts__
```{r}
## Change Value to TS class
gasMonth_ts<-ts(gasMonth,start=c(1986,1),end=c(2016,4),deltat = 1/12)
gasMonth_ts<-gasMonth_ts[,2] # Remove unnecessary column
head(gasMonth_ts)
```
</section>


```{r, echo=FALSE, results='hide'}
## Plot the log of the Gas Price
gasMonth_tsLog<-log(gasMonth_ts)
### Plot the acf for the log of the original Gas Price
##---------------------------------------------------------------
```


## ANALYZE TIME SERIES DATA

<section>
- Plot the data and the ACF
```{r}
## Plot the original Gas Price
ts.plot(gasMonth_ts)
```
</section>

<section>
```{r}
### Plot the acf for the original Gas Price
acf(gasMonth_ts)
```
</section>

<section>
Check stionarity of the log-difference 
```{r, results='hide',echo=FALSE}
## Plot the log difference of the Gas Price
gasMonth_tsLogDiff<-diff(gasMonth_tsLog)
ts.plot(gasMonth_tsLogDiff)
```
</section>

<section>
Check acf of the log-difference 
```{r}
### Plot the acf for the log difference Gas Price
acf(gasMonth_tsLogDiff)
```
</section>


```{r, echo=FALSE,results='hide',warning=FALSE,message=FALSE}
##substract the training data
library(astsa)
library(tseries)
library(fGarch)
gasMonthTrainLogDiff<-ts(gasMonth_tsLogDiff[1:359],start=c(1986,2),end=c(2015,12),deltat = 1/12)

##---------------------------------------------------------------
```

<section>
Based on AIC, the screening result suggests that an ARMA (1,0) is a better.
```{r }
#Write a function to compute AIC of various ARMA orders
AICfn<-function(N,K)
{
  for(i in 1:N)
  {for(j in 1:N)
  {
    print(paste("AIC=",round(AIC(arima(K,order=c(i-1,0,j-1))),0),"p=",i-1,"q=", j-1))
  }}}
AICfn(2,gasMonthTrainLogDiff)
```
</section>


## RESIDUAL ANALYSIS AND GARCH MODEL
<section>
</section>

<section>
Fit ARMA (1,0) and plot the residual plot.
```{r, fig.width=8,fig.height=2.5}
gasMod<-arima(gasMonthTrainLogDiff, order = c(1,0,0));
#Residual analysis
plot(gasMod$residuals,main="Plot of residuals of AR(1) models for the gasMonthTrainLogDiff")

```
</section>

<section>
Use t-test to find the residual is not different from 0.
```{r}
## Use t-test to check the value is significantly different from 0
t.test(gasMod$residuals,mu=0)
```
</section>

<section>
Check the ACF for squared residuals
```{r,results='hide',fig.width=8,fig.height=3.5}
#Residual ACF and PACF plots for squared residuals
acf2(gasMod$residuals^2)
```
</section>


<section>
Use loop to examine ideal GARCH model which GARCH(1,1) is the best
```{r,warning=FALSE}
AICfn2<-function(N,K)
{
  for(i in 1:(N+1))
  { for(j in 1:N)
  {
    print(paste("AIC=",round(AIC(garch(residuals(K),order=c(i,j-1),trace=FALSE)),0),"p=",i,"q=", j-1))
  }}}
AICfn2(2,gasMod)
```
</section>

<section>
Fit time series model for ARMA(1,0)+GARCH(1,1)

```{r,fig.width=8,fig.height=2.5}
gfit_garch<-garchFit(formula=~arma(1,0)+garch(1,1),data = gasMonthTrainLogDiff,
               trace = F,include.mean = F)
u <- gfit_garch@sigma.t
```
</section>

```{r,fig.width=8,fig.height=2.5,echo=FALSE,warning=FALSE, message=FALSE}
arch.fit<-garch(gasMod$residuals,order = c(1,1),trace=F)
arch.fit.res<-resid(arch.fit)[-1]
```

<section>
QQ Plot: ARMA VS ARMA+GARCH

```{r,fig.width=8,fig.height=4,echo=FALSE}
par(mfrow=c(1,2)) # set up
qqnorm(residuals(gasMod))
qqline(residuals(gasMod))
qqnorm(arch.fit.res)
qqline(arch.fit.res)
```
</section>


<section>
```{r,fig.width=8,fig.height=3.5}
## Plot the garch fit 
plot(gasMonthTrainLogDiff, type="l",ylab="GAS Log difference", lwd=1, col="blue")
lines(window(gasMonthTrainLogDiff-2*u), lty=2, col="red")
lines(window(gasMonthTrainLogDiff+2*u), lty=2, col="red")

```
</section>

## FIT TIME SERIES NEURAL NETWORK

<section>
```{r,fig.width=8,fig.height=3}
## Fit Neural Network
library(forecast)
neural.fit<-nnetar(gasMonthTrainLogDiff)
neural.fit.for<-forecast(neural.fit)
plot(neural.fit.for)
```
</section>

## MODEL COMPARISON
```{r, echo=FALSE, results='hide'}
gfit_arima<-arima(gasMonthTrainLogDiff,order = c(1,0,0))
#Predict the fitted value
##obtain predicted value of arima+GARCH model
pre_garch<-predict(gfit_garch,n.ahead=4,trace=FALSE)
##obtain predicted value of pure arima model
pre_arima<-predict(gfit_arima,n.ahead=4,trace=FALSE)
##obtain predicted valeu of neural network model
neural.fit.for<-forecast(neural.fit)
#compare different predicted value
real_value<-gasMonth_tsLogDiff[360:363]
library(pander)

library(Metrics)
##rmse
rmse_garch<-rmse(real_value,pre_garch$meanForecast)
rmse_arima<-rmse(real_value,pre_arima$pred)
rmse_neural<-rmse(real_value,neural.fit.for$mean[1:4])
rmse<-data.frame(rmse_garch,rmse_arima,rmse_neural)
rmse<-c(rmse_garch,rmse_arima,rmse_neural)
##mae
mae_garch<-mae(real_value,pre_garch$meanForecast)
mae_arima<-mae(real_value,pre_arima$pred)
mae_neural<-mae(real_value,neural.fit.for$mean[1:4])
mae<-c(mae_garch,mae_arima,mae_neural)
comparison<-rbind(rmse,mae)
row.names(comparison)<-c("MAE","RMSE")
colnames(comparison)<-c("GARCH","ARMA","NEURAL NETWORK")
```


<section>

Based on MAE and RMSE, the ARMA model has the smallest for these two criteria. 
Therefore, using the ARMA(1,0) to fit the log difference is the best predictor 

for predicting JAN/2016 - APR/2016

```{r,results='asis'}
pandoc.table(comparison)
```
</section>

## THANK YOU {data-background-video="background.mp4"}